Sequential rank agreement methods for comparison of ranked lists

### 2017-04-09 latex/superranker-manuscript.tex
revised simulation section

We present results of a simulation study where we investigated the
small sample properties of our proposed method (sra) and compared it
with the topK method proposed by \citet{hall:schi:2012} with respect
to list depth agreement. In each simulation run, we generated a
dataset with 1000 predictors and 400 observations. The predictors were
drawn independently from a standard Gaussian distribution. The outcome
was drawn from a Gaussian distribution with variance 1 and mean that
depended additively on the first 15 predictors, i.e.,

@@

where $y_i$ is the $i$th response, and $x_{ij}$ is the $j$th predictor
for the $i$th measurement, and $\epsilon_i\sim N(0, 1)$. In each
simulation run we obtained \(L\) ranked lists of the 1000 predictors.
This was done by first drawing \(L\) bootstrap samples (with
replacement) and then in each bootstrap sample ranking the 1000
predictors according to their marginal $t$ test statistics. Thus, the
simulation results compare lists that are obtained with the same
technique on bootstrap versions of the same dataset. We report results
from two sets of simulations (each based on 1000 simulated datasets):
\begin{enumerate}
\item[] Scenario I: \(L=8\) and $q\in\{3,4,5,6,7,8,9,10\}$.
\item[] Scenario II: \(q=5\) and $L\in\{3,5,10,50\}$.
\end{enumerate}
In both scenarios, we summarized the distribution of the estimated
depth of agreement and the average number of different predictors
found at a given depth.
 % Asymptotic results are available for both methods as the number of
 % lists increases, so this will investigate how much the results might
 % change as $L$ increases. For a fixed constant threshold of $q_L=5$ we
 % increase the number of available lists to investigate the sensitivity
 % of the results to $L$.
  % Throughout the simulations we
  % assume a constant, prespecified threshold function, $q_L$, and try
  % different values for the constant. The approach by
  % \citet{hall:schi:2012} uses a window size parameter, $K$, to
  % represent the maximum difference in rankings before they consider an
  % item in two lists to be in agreement.  Because the standard
  % deviation --- which is used in the sequential rank agreement
  % definition of $\hat{A}_L$--- squares its differences, it tends to
  % give more weight to larger rank differences and less weight to
  % smaller differences compared to the mean absolute difference.  To
  % ease comparison between the sequential rank agreement and the
  % \citet{hall:schi:2012} approach we would like to use the same
  % threshold for both methods, and in order to do this we use the
  % median absolute deviance throughout the simulations in lieu of the
  % pooled standard deviation.
 % For all parameter combinations we ran 1000 simulations and report the
 % distribution of the estimated depth of agreement and the average
 % number of different predictors found at a given depth.
The results from Scenario I are shown in the left panel of
Figure~\ref{fig:example99}. The double histograms show the
distribution of the estimated depth of agreement. We see that sra
results in a substantially lower depth of agreement than the topK
method. Also the average numbers of unique predictors (bold numbers in
the double histograms) are markedly smaller for sra. Even larger
differences are found if we used the Euclidean distance instead of the
median absolute distance for the sequential rank agreement (results
not shown). As expected the depth of agreement increases when the
threshold for agreement/window increases.
\begin{figure}[htbp]
   \begin{center}
 \includegraphics[width=.49\textwidth]{Ksim.pdf}
 \includegraphics[width=.49\textwidth]{Lsim}
 \end{center}
 \caption{Left panel: Simulation study showing distribution of
   estimated rank agreements for sra and topK for varying thresholds
   and fixed number of lists $L=8$. The bold numbers are the average
   number of unique predictors in the set found to be in
   agreement. Right panel: Simulation results for varying number of
   lists, and with fixed threshold of $q=5$.}
 \label{fig:example99}
 \end{figure}
 The right panel of Figure~\ref{fig:example99} shows the results from
 Scenario II. The number of lists has little impact but at all depths
 sra is more conservative than topK. As a consequence we also see that
 sra systematically selects fewer predictors in the parts of
 agreement.

>>>>

We present results of a simulation study where we investigated the
small sample properties of the sequential rank agreement and compared
it with the topK method \citep{hall:schi:2012} with respect to list
depth agreement. For the sequential rank agreement we use the function
\texttt{sra} of the \texttt{R} package \texttt{SuperRanker} with the
\texttt{"mad"} option (median absolute distance). To define the depth
of agreement we set a constant threshold function to an integer \(q\)
and report the first crossing point, i.e., the smallest list depth
where sra exceeds \(q\). For topK we use the function
\texttt{j0.multi} which is implemented in the \texttt{R} package
\texttt{TopKLists}. Specifically, we set the tuning parameter
\texttt{v} of \texttt{j0.multi} to the value 6 and the window
parameter \texttt{d} to \(q\) and report the output parameter
\texttt{maxK} as the depth of agreement. Thus, to make this comparison
we assume that the first crossing point of sra and the result of the
topK method measure the same underlying feature.

In each simulation run, we generated a dataset with 1000 predictors
and 400 observations. The predictors were drawn independently from a
standard Gaussian distribution. The outcome was drawn from a Gaussian
distribution with variance 1 and mean that depended additively on the
first 15 predictors, i.e.,

@@

where $y_i$ is the $i$th response, and $x_{ij}$ is the $j$th predictor
for the $i$th measurement, and $\epsilon_i\sim N(0, 1)$. In each
simulation run we obtained \(L\) ranked lists of the 1000 predictors.
This was done by first drawing \(L\) bootstrap samples (with
replacement) and then in each bootstrap sample ranking the 1000
predictors according to their marginal $t$ test statistics. Thus, we
assessed the depth of agreement among lists that are ranked with the
same statistical method on bootstrap versions of the same dataset. We
report simulation results from two scenarios where each is based on
1000 simulated datasets:
\begin{enumerate}
\item[] Scenario I: Fix \(L=8\) and vary $q\in\{3,4,5,6,7,8,9,10\}$.
\item[] Scenario II: Fix \(q=5\) and vary $L\in\{3,5,10,50\}$.
\end{enumerate}
In both scenarios we summarized the distribution of the estimated
depth of agreement and the average number of unique predictors found
in the set of predictors which is selected by the estimated depth of
agreement.
 % Asymptotic results are available for both methods as the number of
 % lists increases, so this will investigate how much the results might
 % change as $L$ increases. For a fixed constant threshold of $q_L=5$ we
 % increase the number of available lists to investigate the sensitivity
 % of the results to $L$.
  % Throughout the simulations we
  % assume a constant, prespecified threshold function, $q_L$, and try
  % different values for the constant. The approach by
  % \citet{hall:schi:2012} uses a window size parameter, $K$, to
  % represent the maximum difference in rankings before they consider an
  % item in two lists to be in agreement.  Because the standard
  % deviation --- which is used in the sequential rank agreement
  % definition of $\hat{A}_L$--- squares its differences, it tends to
  % give more weight to larger rank differences and less weight to
  % smaller differences compared to the mean absolute difference.  To
  % ease comparison between the sequential rank agreement and the
  % \citet{hall:schi:2012} approach we would like to use the same
  % threshold for both methods, and in order to do this we use the
  % median absolute deviance throughout the simulations in lieu of the
  % pooled standard deviation.
 % For all parameter combinations we ran 1000 simulations and report the
 % distribution of the estimated depth of agreement and the average
 % number of different predictors found at a given depth.
The results from Scenario I are shown in the left panel of
Figure~\ref{fig:example99}. The violin plots (with rectangular kernel)
show the distributions of the estimated depths of agreement for both
methods. We see that sra results in a substantially lower depth of
agreement than the topK method. Also the average numbers of unique
predictors (bold numbers inside the plots) are markedly smaller for
sra. Even larger differences were found when we used the Euclidean
distance instead of the median absolute distance for the sequential
rank agreement (results not shown). As expected the depth of agreement
increased when the threshold for agreement/window increased.
\begin{figure}[htbp]
   \begin{center}
 \includegraphics[width=.49\textwidth]{Ksim.pdf}
 \includegraphics[width=.49\textwidth]{Lsim}
 \end{center}
 \caption{Left panel: Simulation study showing distribution of
   estimated rank agreements for sra and topK for varying thresholds
   and fixed number of lists $L=8$. The bold numbers are the average
   number of unique predictors included in the set where the lists
   agree. Right panel: Simulation results for varying number of lists
   and with fixed threshold of $q=5$.}
 \label{fig:example99}
\end{figure}

The right panel of Figure~\ref{fig:example99} shows the results from
Scenario II. The number of lists has little impact on the
results. Generally, sra is more conservative than topK and as a
consequence sra includes fewer predictors in the selected set where
the lists agree.
